---
title: "Syllabus"
reference-location: margin
references:
- id: Frankfurt:2005
  author:
    - family: Frankfurt
      given: Harry G.
  citation-key: Frankfurt:2005
  event-place: Princeton
  issued:
    - year: 2005
  publisher: Princeton University Press
  publisher-place: Princeton
  title: On bullshit
  type: book
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

:::: {.grid .course-details .course-details-smaller}

::: {.g-col-12 .g-col-sm-6 .g-col-md-4}
### Instructor

- [{{< fa user >}}]{.fa-li} [{{< var instructor.name >}}]({{< var instructor.url >}})
- [{{< fa university >}}]{.fa-li} {{< var instructor.office >}}
- [{{< fa envelope >}}]{.fa-li} <a href='mailto:{{< var instructor.email >}}'>{{< var instructor.email >}}</a>
- [{{< iconify ph butterfly-bold size=1.2em title='Bluesky' >}}]{.fa-li} [Bluesky]({{< var instructor.bluesky >}})
- [{{< iconify bi mastodon size=1.1em title='Mastodon' >}}]{.fa-li} [Mastodon]({{< var instructor.mastodon >}})

:::

::: {.g-col-12 .g-col-sm-6 .g-col-md-4}
### Course details

- [{{< fa calendar >}}]{.fa-li} {{< var course.days >}}
- [{{< fa calendar-alt >}}]{.fa-li} {{< var course.dates >}}
- [{{< fa clock >}}]{.fa-li} {{< var course.time >}}
- [{{< fa location-dot >}}]{.fa-li} {{< var course.location >}}
:::

::: {.g-col-12 .g-col-md-4 .contact-policy}
### Contacting me

- [{{< fa calendar-check >}}]{.fa-li} [Schedule an appointment]({{< var instructor.appointment_url >}})
:::

::::


## Course objectives

How do public managers in Mexico, Ghana, and Australia behave differently? Why do some provinces and municipalities perform better than others? How do a country’s political institutions shape its public sector and civil society?

By the end of this course, you will (1) understand the varieties of the public and civil society sectors around the world, (2) see the connections between a country's political regime, governance, bureaucracy, and civil society, (3) make cross-country comparisons of different forms of governing institutions, and (4) use theory and data to argue for specific public policy and management issues. Specifically you'll be able to:

- **Explain** varieties in administrative characteristics and management practices around the world
- **Understand** to what extent different institutional features and management practices influence governance outcomes and public management practices 
- **Apply** the findings of scientific studies to current policy and management issues
- **Critically assess research** about public policy and administration
- **Conduct your own research** on issues in comparative public policy and management 


## Course materials

There is no textbook for this class, and all of the readings are free (!!). You will need to read 2–5 readings (generally policy papers and academic research articles) for each class session, and all of these readings will be available through the GSU library or through iCollege.

::: {.callout-warning}
### Don't buy anything!

**Do not purchase any articles for this class—all readings are freely accessible on iCollege.**
:::


## Attendance and participation

You're expected to come to class each Thursday prepared, having (1) carefully read the material and (2) completed the weekly reading report.

Attendance and participation are crucial to your success in this course. However, we're still in the middle of a global pandemic. If you are sick or are incapable of participating meaningfully in class (e.g. you have stayed up all night and are going to fall asleep in class), please stay home. I will stream our in-person sessions on Thursdays via Webex (and will set up my own camera and microphone system to do it), so if you cannot attend, you can tune in if needed.

GSU has some new process for getting absences excused because of illness, but I don't care about that process. If you're sick, I don't need a doctor's note or anything. If a relative dies and you have to attend a funeral, do it!—don't worry about sending me confirmation or anything. If you have kids and your childcare situation falls through one week, focus on your kids!—again, don't worry about sending me confirmation. You're all adults—I trust you.

If you test positive for COVID-19, hunker down and focus on getting better! Please do not come to class. 


## AI, large language models, and bullshit

I *highly recommend* **not** using ChatGPT or similar large language models (LLMs) in this class. 

I am not opposed to LLMs in many situations. I use [GitHub Copilot](https://github.com/features/copilot) for computer programming-related tasks all the time, and I have ongoing research where we're experimenting with using [Meta's Ollama](https://ollama.com/) to try automatically categorizing thousands of nonprofit mission statements. Using LLMs requires careful skill and attention and practice, and they tend to be useful only in specific limited cases.

I am deeply opposed to LLMs for writing.

Google Docs and Microsoft Word now have built in text-generation tools where you can start writing a sentence and let the machine take over the rest. ChatGPT and other services let you generate multi-paragraph essays with plausible-looking text. **Please do not use these.**

There's a reason most university classes require some sort of writing, like reading reflections, essay questions, and research papers. [The process of writing is actually the process of thinking](https://www.latimes.com/opinion/story/2023-06-20/google-microsoft-chatgpt-ai-writing-assistants-artificial-intelligence):

> Writing is hard because the process of getting something onto the page helps us figure out what we think—about a topic, a problem or an idea. If we turn to AI to do the writing, we're not going to be doing the thinking either. That may not matter if you're writing an email to set up a meeting, but it will matter if you're writing a business plan, a policy statement or a court case. [@Rosenzweig:2023]

Using LLMs and AI to generate a reflection on the week's readings or to generate a policy brief will not help you think through the materials. You can create text and meet the suggested word count and finish the assignment, but the text will be meaningless. There's an official philosophical term for this kind of writing: [bullshit](https://doi.org/10.1007/s10676-024-09775-5) [@HicksHumphriesSlater:2024; @Frankfurt:2005].^[I'm a super straight-laced Mormon and, like, never ever swear or curse, but in this case, the word has a formal philosophical meaning [@Frankfurt:2005], so it doesn't count :)]

Philosophical bullshit is "speech or text produced without concern for its truth" [@HicksHumphriesSlater:2024, p. 2]. Bullshit isn't truth, but it's also not lies (i.e. the opposite of truth). It's text that exists to make the author sound like they know what they're talking about. A bullshitter doesn't care if the text is true or not—truth isn't even part of the equation:

> [A bullshitter] does not reject the authority of the truth, as the liar does, and oppose himself to it. He pays no attention to it at all [@Frankfurt:2005, p. 61].

LLMs and AI systems like ChatGPT, Gemini, Claude, and so on are bullshit machines. That might sound hyperbolic, but at a technological level, that's literally what they do. LLMs use fancy statistics to take an initial prompt or starting phrase and then generate the most plausible words or phrases that are likely to follow. They don't care whether the text it creates is true—they only care about whether it looks plausible. 

This is why these systems will generate citations to articles that don't exist, or generate facts that aren't true, or fail at simple math questions. They aren't meant to tell the truth. Their only goal is to create stuff that looks plausible:

> The problem here isn't that large language models hallucinate, lie, or misrepresent the world in some way. It’s that they are not designed to represent the world at all; instead, they are designed to convey convincing lines of text. [@HicksHumphriesSlater:2024, p. 3]

**Do not replace the important work of writing with AI bullshit slop.** Remember that the point of writing is to help crystalize your thinking. Chugging out words that make it look like you read and understood the articles will not help you learn.

In your weekly reading reports, I want to see good engagement with the readings. I want to see your thinking process. I want to see you make connections between the readings and between real-world events. I don't want to see a bunch of words that look like a human wrote them. That's not useful for future-you. That's not useful for me. That's a waste of time.

I will not spend time trying to guess if your assignments are AI-generated.^[There are tools that purport to be able to identify the percentage of a given text that is AI, but they do not work and result in all sorts of false positives.] If you do turn in AI-produced content, I won't automatically give you a zero. I'll grade your work based on its own merits. I've found that AI-produced content will typically earn a ✓− (50%) or lower on my check-based grading system without me even needing to look for clues that it might have come from an LLM. Remember that text generated by these platforms is philosophical bullshit. Since it has nothing to do with truth, it will not—by definition—earn good grades.


## Pandemic stuff

### Masks and vaccines

::: {.callout-note appearance="simple"}
GSU does not require this and I can't require this but I am allowed to urge it so here's me urging it
:::

Please get the COVID-19 vaccination and a booster shot. It is free. It saves lives. (I am fully vaccinated and repeatedly-boosted.)

GSU and the University System of Georgia do not have a mask mandate for students or faculty. However, I personally will be wearing a mask when [the CDC's wastewater surveillance ratings for Georgia](https://www.cdc.gov/nwss/rv/COVID19-currentlevels.html) are moderate or above.

**I strongly recommend/urge/encourage you to wear a mask in class on Thursdays** regardless of your vaccination status. I will place a box of disposable masks by the door if you would like one. 

Again, GSU does not require either vaccines or masks, and if you aren't vaccinated or don't wear a mask, there are no penalties.

### Learning during a pandemic

Life still sucks right now. None of us is really okay. **We're all just pretending.**

You most likely know people who have lost their jobs, have been hospitalized, or have even died (I myself know people in all those categories). You all have increased (or possibly decreased) work responsibilities and increased family care responsibilities—you might be caring for extra people (young and/or old!) right now, and you are likely facing uncertain job prospects (or have been laid off!).

**I'm fully committed to making sure that you learn everything you were hoping to learn from this class!** I will make whatever accommodations I can to help you finish your problem sets, do well on your projects, and learn and understand the class material. Under ordinary conditions, I am flexible and lenient with grading and course expectations when students face difficult challenges. Under pandemic conditions, that flexibility and leniency is intensified.

If you tell me you're having trouble, I will not judge you or think less of you. I hope you'll extend me the same grace.

You *never* owe me personal information about your health (mental or physical). You are *always* welcome to talk to me about things that you're going through, though. If I can't help you, I usually know somebody who can.

If you need extra help, or if you need more time with something, or if you feel like you're behind or not understanding everything, **do not suffer in silence!** Talk to me! I will work with you. **I promise.**

*Please* sign up for a time to meet with me during student hours at <https://calendly.com/andrewheiss/>. I'm also available through e-mail.

I want you to learn lots of things from this class, but I primarily want you to stay healthy, balanced, and grounded during this crisis.



## Course policies

**Be nice. Be honest. Don't cheat.**

We will also follow [Georgia State's Code of Conduct](https://codeofconduct.gsu.edu/).

This syllabus reflects a plan for the semester. Deviations may become necessary as the semester progresses.

### Student hours

Please watch this video:

<iframe src="https://player.vimeo.com/video/270014784" width="640" height="360" frameborder="0" allow="autoplay; fullscreen" allowfullscreen style="display: block; margin: 0 auto 1rem;"></iframe>

Student hours are set times dedicated to all of you (most professors call these "office hours"; I don't^[There's fairly widespread misunderstanding about what office hours actually are! [Many students often think that they are the times I *shouldn't* be disturbed](https://www.chronicle.com/article/Can-This-Man-Change-How-Elite/245714/), which is the exact opposite of what they're for!]). This means that I will be ~~in my office~~ at home (wistfully) waiting for you to ~~come by~~ talk to me remotely with whatever questions you have. This is the best and easiest way to find me and the best chance for discussing class material and concerns.

Because of the pandemic, it is easiest to meet with me online via Webex for student hours. [Make an appointment with me here](https://calendly.com/andrewheiss/), and then use this link to talk to me during student hours: <https://gsumeetings.webex.com/meet/aheiss>. You can also find me through e-mail.

### Late work

You will lose 2 points per day for each day an assignment is late. This is designed to not be a huge penalty (2 days late = 16/20 points on a reading report that gets a ✓), but instead works a commitment device to help you stay on schedule.

### Counseling and Psychological Services (CPS)

Life at GSU can be complicated and challenging (especially during a pandemic!). You might feel overwhelmed, experience anxiety or depression, or struggle with relationships or family responsibilities. [Counseling and Psychological Services (CPS)](https://education.gsu.edu/cps/) provides free, *confidential* support for students who are struggling with mental health and emotional challenges. The CPS office is staffed by professional psychologists who are attuned to the needs of all types of college and professional students. Please do not hesitate to contact CPS for assistance—getting help is a smart and courageous thing to do.

### Basic needs security

If you have difficulty affording groceries or accessing sufficient food to eat every day, or if you lack a safe and stable place to live, and you believe this may affect your performance in this course, please contact the [Dean of Students](https://deanofstudents.gsu.edu/) for support. They can provide a host of services including free groceries from the [Panther Pantry](https://nutrition.gsu.edu/panther-pantry/) and assisting with homelessness with the [Embark Network](https://deanofstudents.gsu.edu/student-assistance/embark/). Additionally, please talk to me if you are comfortable in doing so. This will enable me to provide any resources that I might possess.

### Lauren's Promise

**I will listen and believe you if someone is threatening you.**

Lauren McCluskey, a 21-year-old honors student athlete, [was murdered on October 22, 2018 by a man she briefly dated on the University of Utah campus](https://www.sltrib.com/opinion/commentary/2019/02/10/commentary-failing-lauren/). We must all take action to ensure that this never happens again.

If you are in immediate danger, call 911 or GSU police (404-413-3333).

If you are experiencing sexual assault, domestic violence, or stalking, please report it to me and I will connect you to resources or call [GSU's Counseling and Psychological Services](https://counselingcenter.gsu.edu/crisis-services/concern-self/immediate-help/) (404-413-1640).

Any form of sexual harassment or violence will not be excused or tolerated at Georgia State. GSU has instituted procedures to respond to violations of these laws and standards, programs aimed at the prevention of such conduct, and intervention on behalf of the victims. Georgia State University Police officers will treat victims of sexual assault, domestic violence, and stalking with respect and dignity. Advocates on campus and in the community can help with victims' physical and emotional health, reporting options, and academic concerns.

### Academic honesty

Violation of [GSU's Policy on Academic Honesty](https://deanofstudents.gsu.edu/faculty-staff-resources/academic-honesty/) will result in an F in the course and possible disciplinary action.^[So seriously, just don't cheat or plagiarize!] All violations will be formally reported to the Dean of Students.

### Special needs

Students who wish to request accommodation for a disability may do so by registering with the [Office of Disability Services](https://disability.gsu.edu/). Students may only be accommodated upon issuance by the Office of Disability Services of a signed [Accommodation Plan](https://disability.gsu.edu/services/how-to-register/) and are responsible for providing a copy of that plan to instructors of all classes in which accommodations are sought.

Students with special needs should then make an appointment with me during the first week of class to discuss any accommodations that need to be made.


## Global Scholar Distinction

This is a Global Scholar Distinction course. The GSD can be earned by students in any major and demonstrates a readiness to learn and lead with a global perspective. The GSD leads to a badge at graduation that can be added to your LinkedIn profile and other online profiles. [Learn more here](https://international.gsu.edu/initiatives/global-scholar-distinction/).

For more information about study abroad, the free passport initiative, virtual exchange classes and more through the GSU office of International Initiatives, visit <https://international.gsu.edu/>


## Assignments and grades

You can find descriptions for all the assignments on the [assignments page](/assignment/).

```{r assignments-grades, include=FALSE}
assignments <- tribble(
  ~Points, ~Assignment,
  13 * 20, "Reading reports (13 × 20)",
  40,      "Presentation and discussion",
  125,     "Data-based essay",
  125,     "Theory in real life essay",
  250,     "Policy brief"
) %>%
  mutate(Percent = Points / sum(Points),
         Percent = scales::percent(Percent))

grading <- tribble(
  ~Grade, ~Range, ~Grade1, ~Range1,
  "A",  "93–100%", "C",  "73–76%",
  "A−", "90–92%", "C−", "70–72%",
  "B+", "87–89%", "D+", "67–69%",
  "B",  "83–86%", "D",  "63–66%",
  "B−", "80–82%", "D−", "60–62%",
  "C+", "77–79%", "F",  "< 60%"
)
```

::: {.centered-table}
```{r show-assignments-table, echo=FALSE, results="asis"}
assignments_total <- tibble(Assignment = "Total",
                            Points = sum(assignments$Points),
                            Percent = NA)

bind_rows(assignments, assignments_total) %>%
  select(Assignment, Points, Percent) %>%
  pandoc.table(justify = "lcc", missing = "—")
```
:::

::: {.centered-table}
```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lclc",
               col.names = c("Grade", "Range", "Grade", "Range"))
```
:::


## Obscure animals

Once you have read this entire syllabus and [the assignments page](/assignment/), please [click here](mailto:aheiss@gsu.edu) and e-mail me a picture of the most obscure animal you can think of.
